
\section{List of examples}

\gls{persn} markers are set in \textbf{bold} for ease of access.


\subsection{Afroasiatic languages}


\subsubsection{Hausa (haus1257), West Chadic}

\pex
\a \begingl
\gla \textbf{m\={u}} Háus\textgravemacron{a}w\={a}//
\glb we Hausa//
\glft `we Hausa'\\\citep[371]{newman2000}//
\endgl
\a \begingl
\gla \textbf{sh\={\textsci}} wannàn m\={a}làm\={\textsci}//
\glb he \Dem{}.1 teacher//
\glft `he (this) teacher'\\{\citep[after][371]{newman2000}}//
\endgl
\a
\begingl
\gla \textbf{m\={u}} m\textgravemacron{a}làman-nàn//
\glb we teacher-\Dem.\Prox{}//
\glft `we these teachers'\\\citep[155]{newman2000}//
\endgl
\xe

See \citet[63, 155, 370f.]{newman2000} and also \citet[330f.]{jaggar2001} for further examples.




\subsubsection{Mupun (mwag1236), West Chadic}
 
\ex
\begingl
\gla \textbf{war} manaja n\textschwa//
\glb 3\F{} manager \Def{}//
\glft `she, the manager'\\\citep[after][172, (154)]{frajzyngier1993}//
\endgl
\xe
 
\subsubsection{Gorwaa (goro1270), Southern Cushitic}

\ex
\begingl
\gla \textbf{atén} oo hhawató//
\glb Pro1\Pl{} \Anaph.\M{} men.\Lnk.\M{}//
\glft `we men'\\\citep[after][163, (2.205)]{harvey2018}//
\endgl
\xe

\subsubsection{Kambaata (kamb1316), East Cushitic}

\ex 
\begingl
\gla \textbf{na'óot} Kambáat-u//
\glb 1\Pl.\Nom{} Kambaata-\M.\Nom{}//
\glft `we Kambaata (people)' \\\citep[335, (1097)]{treis2008}//
\endgl
\xe

\subsubsection{Cairene Egyptian Colloquial Arabic (egyp1253), Semitic}

\ex
\begingl
\gla \textbf{\textglotstop{}intu} \textglotstop{}it[-]talamza ti\textcrh{}ibbu \textglotstop{}illi\textrevglotstop{}b//
\glb you.\Pl{} [\Def-]students 2\Pl{}.like playing//
\glft `You students like playing.'\\\citep[80, (533)]{garygamal1982}//
\endgl
\xe

See \citealp[78; 80]{garygamal1982}.

\subsubsection{Gulf Arabic (gulf1241), Semitic}


\ex
\begingl
\gla \textbf{iHna} T-Tullaab ma nigdar nigbal haadha l-qaraar//
\glb we \Def-students not 1\Pl.be.able 1\Pl.accept this \Def-decision//
\glft `We students cannot accept this decision'\\\citep[165, (845)]{holes1990}//
\endgl
\xe


See \citealp[162, 165]{holes1990}.

\subsubsection{Maltese (malt1254), Semitic}

\ex
\begingl
\gla {}[\textbf{Intom} il-\textcrh{}addiema] g\textcrh{}andkom ting\textcrh{}aqdu//
\glb you \Def-workers have.2\Pl{} unite.2\Pl{}//
\glft `You workmen should unite together.' \\{\citep[202, (915)]{borgazzopardialexander1997}}//
\endgl
\xe 


%%%%%%%%%%%%%%%
\subsection{Australian languages}

\subsubsection{Mangarrayi (mang1381), Mangarrayi}

\ex
\begingl
\gla \textbf{\ng{}j\d{l}a} malam-ga\d{l}a ga-\ng{}i\d{l}a-\d{n}i \d{n}a-wa\d{l}ayjñin-gan//
\glb 1\Pl.\Excl.\Nom{} man-\Pl.\Nom{} 3-1\Pl.\Excl-sit NLoc-shade//
\glft `We men are sitting in the shade.'\\\citep[103]{merlan1989}//
\endgl
\xe

See \citealp[103; 203]{merlan1989} for discussion.


\newpage
\subsubsection{Diyari (pirl1241), Central Pama-Nyungan}


\pex 
\a \begingl
\gla \textbf{ngayani} waka-li thananha nhayi-yi//
\glb 1\Pl.\Excl.\Nom{} small-\Erg{} 3\Pl.\Acc{} see-\Prs{}//
\glft `We small (ones) watch them'\\\citep[102, (106)]{austin2013}//
\endgl
\a 
\begingl
\gla ngarda-nhi thana-li nhinha yakalka-yi yaru-ya wardayari-lha \textbf{yini} pinarru//
\glb then-\Loc{} 3\Pl.\Erg{} 3\Sg.\Nf.\Acc{} ask-\Prs{} like-\Dem.\Prox{} where-\Char{} 2\Sg.\Nom{} old.man.\Nom//
\glft `Then they asked him the following: ``Where are you from, old man?'' '\\\citep[119, (175)]{austin2013}//
\endgl
\a\begingl
\gla \textbf{nhani} mankarra thurara-yi thana-ngu-ya kinthala-nhi//
\glb 3\Sg.\F.\Nom{} girl.\Nom{} lie-\Prs{} 3\Pl-\Loc-\Prox{} dog-\Loc{}//
\glft `The girl is lying with those dogs'\\\citep[150, (311)]{austin2013}//
\endgl
\xe


For discussion and more examples see \citet[97f.]{austin1981} and \citep[100--103, 119, 150]{austin2013}.

\subsubsection{Warlpiri (warl1254), Western Pama-Nyungan}

\pex \a 
\begingl
\gla \ng{}arka \textbf{njanu\ng{}u} ka pu\d{l}a-mi//
\glb man 3 \Prs{} shout-\Npst{}//
\glft `The aforementioned man is shouting'\\\citep[316, (22)]{hale1973}//
\endgl
\a
\begingl
\gla yapa \textbf{\ng{}atju}//
\glb person 1\Sg{}//
\glft `I person'\\\citep[317]{hale1973}//
\endgl
\xe

See also \citealp[70]{reece1970} and especially \citealp[316f.]{hale1973} for further discussion.


\newpage

\subsubsection{Pitjantjatjara (pitj1243), Western Pama-Nyungan}

\pex
\a
\begingl
\gla Minyma \textbf{palu\underline{r}u} ngayu-nya nya-ngu//
\glb woman 3\Sg{}.\Nom{} 1\Sg{}.\Acc{} see-\Pst{}//
\glft `The woman saw me.' \\{\citep[31, (110)]{bowe1990}}//
\endgl
\a 
\begingl
\gla \textbf{Palu\underline{r}u} wati nyara wa\underline{r}a-ngku mutaka palya-nu//
\glb 3\Sg{}.\Nom{} man distant tall-\Erg{} car fix-\Pst{}//
\glft `The tall man over there (in contrast to the other one) fixed the car.' \\{\citep[34, (114)]{bowe1990}}//
\endgl
\xe

See \citet[49--51]{bowe1990} for discussion.

\subsubsection{Guugu Yimidhirr (gugu1255), Northern Pama-Nyungan}

\pex
\a
\begingl
\gla \textbf{Nyulu} nhayun waarigan gada-y waanggu=wunaarna-y.//
\glb 3\Sg.\Nom{} that.\Abs{} moon.\Abs{} come-\Pst{} sleep=lie+\Redup-\Pst//
\glft `[Then] the Moon came and lay down to sleep.'\\\citep[157, (423)]{haviland1979}//
\endgl
\a
\begingl
\gla Bidha \textbf{nyulu} biini.//
\glb child.\Abs{} 3\Sg.\Nom{} die.\Pst{}//
\glft `The child died.'\\\citep[157, (424a)]{haviland1979}//
\endgl
\xe

For general description suggesting that ``the norm arrangement for an NP that refers to a human'' involves a prenominal \gls{apc} see \citet[104]{haviland1979}.

\subsubsection{Kuku Yalanji (kuku1273), Northern Pama-Nyungan}

\pex
\a 
\begingl
\gla \textbf{nyulu} jalbu//
\glb 3\Sg{} woman//
\glft `\textbf{the} woman'\\{\citep[after][202; gloss extrapolated]{patz2002}}//
\endgl
\a
\begingl
\gla \textbf{Yurra} karrkay dunga-y bana mana!//
\glb 2\Pl{}.\Nom{}(\Sarg{}) child.\Abs{}(\Sarg{}) go-\Imp{} water.\Abs{}(\Obj{}) get.\Imp{}//
\glft `You children go and get water!' \\{\citep[after][203, (620)]{patz2002}}//
\endgl
\xe

\pex
\a 
\begingl
\gla Pastor \textbf{nyulu}\ldots//
\glb pastor 3\Sg{}//
\glft `Pastor, he\ldots'\\{\citep[after][202, (611); gloss extrapolated]{patz2002}}//
\endgl
\a
\begingl
\gla Ngayu babi wilbuman yindu ngamu \textbf{nganjin} dunga-ri-ny mayi baka-nka.//
\glb 1\Sg{}.\Nom{}(\Sarg{}) father's.mother.\Abs{}(\Sarg) old.woman.\Abs{}(\Sarg{}) other.\Abs{}(\Sarg{}) mother.\Abs{}(\Sarg{}) 1\Pl{}.\Excl{}.\Nom{}(\Sarg{}) go-\Pl{}-\Pst{} food.\Abs{}(\Obj{}) dig-\Purp//
\glft `I, grandmother, another old woman and mother, we went out to dig for food (yams).' \\{\citep[after][203, (618)]{patz2002}}//
\endgl
\xe

See \citet[120f., 202f.]{patz2002}.

\subsubsection{Kayardild (kaya1319), Tangkic}

\ex
\begingl
\gla \textbf{niya} jungarra dangkaa//
\glb he big man//
\glft `the big man'\\\citep[239]{evans1995}//
\endgl
\xe

See also \citet[239; 251]{evans1995} and \citet[141]{round2013}.


%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{Austronesian languages}

\subsubsection{Malagasy (mala1537), Barito}

\ex
\begingl
\gla \textbf{izaho} vehivavy//
\glb 1\Sg.\Def{} woman//
\glft `I woman'\\\citep[411, (6a)]{paultravis2019}//
\endgl
\xe

For further discussion of some complex variation patterns see \citet{paultravis2019}.
 
 
 \newpage
\subsubsection{Indonesian (stin1234), Malayo-Sumbawan}
	
\ex \begingl
\gla \textbf{kami}, bangsa Indonesia//
\glb we people Indonesia//
\glft `we, the people of Indonesia'\\\citep[170]{sneddon1996}//
\endgl
\xe

\subsubsection{Papuan Malay (papu1250), Malayo-Sumbawan}

\ex
\begingl
\gla de blang, a, \textbf{om} \textbf{ko} \textbf{ini} tra liat\ldots//
\glb 3\Sg{} say ah! uncle 2\Sg{} \Dem.\Prox{} \Neg{} see//
\glft `he said, ``ah, \textbf{you uncle there} didn't see\ldots'' '\\\citep[353, (66)]{kluge2017}//
\endgl
\xe

See \citet[ch. 6.2]{kluge2017} for detailed discussion and more examples.

\subsubsection{Loniu (loni1238), Oceanic}
	

\pex
\a \begingl
\gla \textbf{s\textepsilon{}h} pihin s\textepsilon{}h \v{c}ani \textbf{uw\textepsilon{}h} kaman uw\textepsilon{}h w\textepsilon{}\v{c}\textepsilon{} ake//
\glb 3\Pl{} woman 3\Pl{} clear 1\Pl.\Excl{} man 1\Pl.\Excl{} cut.down tree//
\glft `The women clear, we men cut down the trees.'\\\citep[90, (6)]{hamel1994}//
\endgl
\a 
\begingl
\gla \textbf{iy} amat iy\textopeno{}//
\glb 3\Sg{} man \Dem//
\glft `this man'\\\citep[100, (89)]{hamel1994}//
\endgl
\xe

See \citet[sec. 4.2.1]{hamel1994} for discussion.

\subsubsection{Windesi Wamesa (wame1241), Oceanic}

\ex
\begingl
\gla sinitu=pa-\textbf{tata}//
\glb person=\Det{}-1\Pl{}.\Incl{}//
\glft `we people' \\{\citep[144, (3.46)]{gasser2014}}//
\endgl
\xe

\subsubsection{Maori (maor1246), Oceanic}

\ex 
\begingl
\gla E kaha rawa atu [\textbf{maatou} ngaa  kaiako naa] ki te pata\textperiodcentered{}patai//
\glb \Tam{} strong very away 1\Pl.\Excl{} the.\Pl{} teacher \Dem.2 to the \Redup{}\textperiodcentered{}ask//
\glft `We teachers ask a lot of questions.'\\\citep[373, (1673)]{bauer1993}//
\endgl
\xe

For discussion and more examples see \citealp[368, 373]{bauer1993} and \citealp[262f.]{bauer1997}.

\subsubsection{Tuvaluan (tuva1244), Oceanic}

\pex \a
\begingl
\gla \textbf{Au} ttino poto koo leva ne iloa nee au mea kolaa faatoaa iloa nee koe ttagata valea.//
\glb I the+person intelligent \Pfv{} {} {} know \Erg{} I thing those just know \Erg{} you the+man stupid//
\glft `I, an intelligent person, have long known what you, stupid man, are just discovering.'//
\endgl
\a\begingl
\gla \textbf{Taatou} tino Tuuvalu e see tau ki meakkai kolaa.//
\glb we.\Incl{} person Tuvalu \Npst{} \Neg{} befit to food those//
\glft `We Tuvaluans are not accustomed to that [type of] food.'\\\citep[393, (2018/2019)]{besnier2000}//
\endgl
\xe

See \citealp[392f.]{besnier2000}.

\subsubsection{Kwaio (kwai1243), Oceanic}

\pex
\a \begingl
\gla \textbf{'a-gauru-a} ta'a i 'Ai'eda//
\glb  \Fpron.3\Tri-? people \Loc{} 'Ai'eda//
\glft `those 'Ai'eda people'//
\endgl
\a
\begingl
\gla fa-\textbf{meru-a} ta'a geni//
\glb for-\Spron.1\Tri.\Excl-? people female//
\glft `for us women' \\\citep[after][104]{keesing1985}//
\endgl
\xe

The status of the \emph{-a} intervening between pronoun and noun is not clear, see \citet[25]{hoehn2020ThirdGap} for speculation that this might be a reduced article.

\subsubsection{Arop-Lokep (arop1243), Oceanic}

\ex
\begingl
\gla \textbf{am} garup ke Bok//
\glb 1\Excl.\Pl{} female.one of Bok//
\glft `we women of Bok village'\\\citep[255]{djernes2002}//
\endgl
\xe


\subsubsection{Cheke Holo (chek1238), Oceanic}

\pex 
\a \begingl
\gla \textbf{Tahati} naikno \={g}re e kmana pui~puhi=da//
\glb 1\Pl.\Incl{} people \Dem.\Prox.\Pl{} \Emph{} lot.of \Dur~way=1\Pl.\Poss{}//
\glft `We people have had many problems.'\\\citep[165, (591)]{boswell2018}//
\endgl
\a \begingl
\gla \textbf{Gotilo} Honiara fa-le~lehe egu//
\glb 2\Pl{} Honiara \Caus-\Dur~die like.that//
\glft `You people of Honiara kill [people], like that.'\\\citep[99, (301)]{boswell2018}//
\endgl
\xe

%See \citealp[165]{boswell2018}.

\subsubsection{Hoava (hoav1238), Oceanic}


\ex \begingl
\gla Maki lavati sa pa Solomone, gi ta-\underline{n}ani \textbf{gita} nikana hupa.//
\glb \Neg{}.\Imp{} big 3\Sg{} \Loc{} Solomons and \Pass-eat 1\Pl.\Incl{} man black//
\glft `Let it [a monkey] not grow big in the Solomons, and we black men be bitten.'\\\citep[after][48, (72a)]{davis2003}//
\endgl
\xe

See \citealp{palmer2017} for detailed discussion.

\subsubsection{Kokota (koko1269), Oceanic}

\pex
\a 
\begingl
\gla \textbf{gai} nakoni zuzufra//
\glb we.\Excl{} person black//
\glft `we black people'\\\citep[95, (3.70a)]{palmer2008}//
\endgl
\a 
\begingl
\gla ka \textbf{gai} ira nakoni zuzufra tana nogoi naito tahi ke a\={g}e=u=ni=u//
\glb \Loc{} we.\Incl{} the.\Pl{} person black then \Voc{} devil sea \Pfv{} go=be.thus=3\Sg.\Obj=\Cnt//
\glft `With us black people, then, man!, `sea devil' is what it's called.' \\\citep[123, (4.1b)]{palmer2008}//
\endgl
\xe

See \citet[68, 95, 116, 119, 123, 131, 137, 163, 242, 300, 305, 327, 399, 414]{palmer2008} for various examples.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{Dravidian languages}


\subsubsection{Kannada (nucl1305), Dravidian}

\ex \begingl
\gla \textbf{na:vu} bra:hmaNaru vidyeyannu mareyaba:radu.//
\glb we Brahmin.\Pl{} education.\Acc{} forget.\Inf.\Proh//
\glft `We Brahmins shouldn't forget education.'\\\citep[209, (755)]{sridhar1990}//
\endgl
\xe

See \citet[205, 208f.]{sridhar1990}.

\subsubsection{Malayalam (mala1464), Dravidian}

\ex \begingl
\gla \textbf{\textltailn{}a\ng\ng{}a\textrtaill{}} intyakkaar adhikavum adhikavum sasyabhuukkuka\textrtaill{} aa\textrtailn{}\textschwa{}//
\glb we.\Excl{} Indians mostly vegetarian.\Pl{} be.\Prs{}//
\glft `We Indians are mostly vegetarians.'\\\citep[262, (1340)]{asherkumari1997}//
\endgl
\xe 

See  \citealp[262f.]{asherkumari1997} for more examples.

\subsubsection{Tamil (tami1289), Dravidian}

\ex \begingl
\gla \textbf{naa\ng{}ka} i\ng{}lii\c{s}kaara\ng{}kellaam kaaramaana vastu caappi\textrtailt{}aratille//
\glb we English.person.\Pl.all spicy thing eat.\Prs.\Nom{}.\Neg//
\glft `We English do not eat hot things.'\\\citep[146]{asher1985}//
\endgl
\xe 

See \citet[142, 146]{asher1985}.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{Indo-European languages}


\subsubsection{Welsh (wels1247), Celtic}

\ex \begingl
\gla \textbf{ni} fyfyrwyr//
\glb we \Softmut.student.\Pl{}//
\glft 'we students'\\(David Willis, personal communication)//
\endgl
\xe

\subsubsection{Danish (dani1285), Germanic}

\ex \begingl
\gla lad \textbf{os} voksne snakke i fred//
\glb let.\Imp{} 1.\Pl.\Acc{} adults talk.\Inf{} in peace//
\glft `Let us adults talk in peace.’\\\citep[29, (32b)]{schroeter2021}//
\endgl
\xe

See \citet{schroeter2021} for detailed discussion and \citet{johannessen2008} for psychologically distant demonstratives, corresponding to adnominal third person singular pronouns. 

\subsubsection{Icelandic (icel1247), Germanic}

\ex
\begingl
\gla við/ þið Íslendigar//
\glb we you.\Pl{} Icelanders//
\glft `we/ you Icelanders'\\\citep[11, (16a)]{sigurdssonwood2020}//
\endgl
\xe

See \citet[11f.]{sigurdssonwood2020} for further examples and \citet{johannessen2008} for psychologically distant demonstratives, corresponding to adnominal third person singular pronouns. 

\subsubsection{Norwegian (norw1258), Germanic}

\ex 
\begingl
\gla til \textbf{oss} (to) (gaml-e) professor-a-ne//
\glb to us two old-\Def{} professor-\Pl-\Def{}// 
\glft `to us (two) (old) professors' \\\citep[129, (4.36)]{julien2005}//
\endgl
\xe

See also \citet{johannessen2008} for psychologically distant demonstratives, corresponding to adnominal third person singular pronouns. 

\subsubsection{Swedish (swed1254), Germanic}

\ex 
\begingl
\gla \textbf{vi} student-er//
\glb we student-\Pl{}//
\glft `we students'\\\citep[128, fn. 18, (ia)]{julien2005}//
\endgl
\xe

See also \citet{johannessen2008} for psychologically distant demonstratives, corresponding to adnominal third person singular pronouns. 

\subsubsection{Dutch (dutc1256), Germanic}

\ex
\begingl
\gla \textbf{Wij}/ \textbf{jullie} taalkundigen denken te veel na.//
\glb we you.\Pl{} linguists think too much \Prtcl{}//
\glft `We/you linguists think too much.'\\\citep[52, (23a)]{corver2008vocative}//
\endgl
\xe 


\subsubsection{English (stan1293), Germanic}

\ex \textbf{you} linguists\xe

See among others: \citet{postal1969, delormedougherty1972, sommerstein1972, pesetsky1978, keizer2016}.

\subsubsection{German (stan1295), Germanic}

\ex
\begingl
\gla Wenn noch nicht einmal [\textbf{du} Linguist] die + neue Rechtschreibung beherrschst\ldots//
\glb if still \Neg{} \Prtcl{} you.\Sg{} linguist \Det.\Acc.\Sg{} new orthography command.2\Sg{}//
\glft `If not even you linguist have a command of the new orthography rules\ldots' \\{\citep[after][100, (36)]{rauh2004}}//
\endgl
\xe

See among others: \citet[ch. 6]{lawrenz1993}, \citet{rauh2003, rauh2004} and \citet{roehrs2005}.

\subsubsection{Standard Modern Greek (mode1248), Greek}

\ex 
\begingl
\gla \textbf{Emis} i ghlosoloji piname.//
\glb we \Det.\Pl{} linguists are.hungry.1\Pl{}//
\glft `We linguists are starving/hungry.'\\\citep[after][114, (12c)]{lekakouszendroi2012}//
\endgl
\xe 

See \citet{stavrou1995} as well as \citet[chs. 1/2]{choi2014phd} and \citet[sec. 5]{hoehn2015unagr}.

\subsubsection{Calabrian Greek/Greko (apul1236), Greek}

\ex
\begingl
\gla \textbf{Emì} ta pedìa den pìnnome kafè.//
\glb we \Det.\N.\Pl{} children.\N{} \Neg{} drink.1\Pl{} coffee//
\glft `We children don't drink coffee.'\\\citep[after][274, (17)]{hoehnetalICGLcalabria}//
\endgl
\xe

See \citet{hoehnetalICGLcalabria} for further discussion.

\subsubsection{Kashmiri (kash1277), Indic}

\ex 
\begingl
\gla \textbf{\textschwa{}s'} k\textschwa{}:shir' chi k\textschwa{}:r'gar.//
\glb we Kashmiris are artisans//
\glft `We Kashmiris are artisans.'\\\citep[200, (8)]{walikoul1997}//
\endgl
\xe

See \citet[200]{walikoul1997}.

\subsubsection{Marathi (mara1378), Indic}

\ex \begingl
\gla \textbf{\={a}mh\={i}} b\={a}yak\={a} nehm\={i}ts \={a}ply\={a} kutumb\={ã}n s\={a}\d{t}h\={i} khapto//
\glb we women always-\Emph{} \Refl{} families for {work.hard-1\Pl}//
\glft `We women always work hard for the sake of our families.'\\\citep[386, (1141)]{pandharipande1997}//
\endgl
\xe

See \citet[381, 386]{pandharipande1997} for more examples.


\subsubsection{Punjabi (panj1256), Indic}

\ex
\begingl
\gla \textbf{as\={ii}} pañjaabii garam mijaaj de ãã.//
\glb we Punjabi hot nature \Gen.\M.\Pl{} are//
\glft `We Punjabis are hot blooded/tempered.'\\\citep[228, (703)]{bhatia1993}//
\endgl
\xe

See \citet[228]{bhatia1993}.

\subsubsection{Persian (west2369), Iranian}

\ex 
\begingl
\gla \textbf{ma} irani-a//
\glb we Iranian-\Pl//
\glft `we Iranians'\\\citep[212, (356)]{mahootian1997}//
\endgl
\xe

See \citet[209, 212]{mahootian1997}.

\subsubsection{Aromanian (arom1237), Romance}

\ex \begingl
\gla \textbf{noi} pikurar-li adrem pini.//
\glb we shepherd-\Def.\Pl{} baled bread//
\glft `We shepherds baked bread.'\\(elicited)//
\endgl
\xe

See also \citet[560]{hoehn2015unagr}. 

\subsubsection{Romanian (roma1327), Romance}

\ex \begingl
\gla \textbf{Voi} avoca\cb{t}ii v\u{a} ap\u{a}ra\cb{t}i clien\cb{t}ii.//
\glb you.\Pl{} lawyers-\Def{} \Cl.2\Pl{} defend cuzstomers-\Def{}//
\glft `You lawyers defend your clients.'\\\citep[10, (20a)]{cornilescunicolae2014}//
\endgl
\xe

For further examples see \citet[255, 258]{mallinson1986} and \citet[6, 10, 20f.]{cornilescunicolae2014}.

\subsubsection{Catalan (stan1289), Romance}

\ex \begingl
\gla \textbf{nosaltres} els bombers//
\glb we \Det.\Pl{} firemen//
\glft `we firemen'\\\citep[290]{hualde1992catalan}//
\endgl
\xe 

See also \citet[287, 290]{hualde1992catalan} and \citet[560]{hoehn2015unagr}.

\subsubsection{Galician (gali1258), Romance}

\ex
\begingl
\gla \textbf{nos} os estudantes//
\glb we \Det.\Pl{} students//
\glft `we students'\\\citep[560, (36a)]{hoehn2015unagr}//
\endgl
\xe

See \citealp[560]{hoehn2015unagr}, also \citet[152, 301]{alvarezetal1986}.

\subsubsection{European Portuguese (port1283), Romance}

\ex 
\begingl
\gla \textbf{Nós} portugueses bebemos bom café.//
\glb we Portuguese drink.1\Pl{} good coffee//
\glft `We Portuguese drink good coffee.'\\\citep[555]{hoehn2015unagr}//
\endgl
\xe

\subsubsection{Spanish (stan1288), Romance}

\ex \begingl
\gla \textbf{Nosotros} los lingüistas somos listos.//
\glb we \Det.\Pl{} linguists be.1\Pl{} smart//
\glft `We linguists are smart.'\\\citep[210, (38a)]{choi2014phd}//
\endgl
\xe

See \citet[210f.]{choi2014phd} and \citet[560]{hoehn2015unagr} as well as \citet[145]{debruyne1995}.

\newpage

\subsubsection{Italian (ital1282), Romance}

\ex \begingl
\gla \textbf{noi}/ \textbf{voi} linguisti//
\glb we you.\Pl{} linguists//
\glft `we/you linguists'\\\citep[202, (21a)]{cardinaletti1994}//
\endgl
\xe

See \citet[202f.]{cardinaletti1994}, \citet[559]{hoehn2015unagr} and also \citet{hoehnetal2016CalabrUnagr,hoehnetalICGLcalabria}.

\subsubsection{Northern Calabrese (nort2612), Romance}

\ex \begingl
\gla \textbf{Nua} i quatrarə iucamə i cartə.//
\glb we \Det.\Pl{} children play.1\Pl{} \Det.\Pl{} cards//
\glft `We children play cards.'\\\citep[276, (23b)]{hoehnetalICGLcalabria}//
\endgl
\xe

See similarly \citet[142]{hoehnetal2016CalabrUnagr}.

\subsubsection{Southern Calabrese (sout2616), Romance}

\ex \begingl
\gla \textbf{Nui} i figghioli iocamu e carti.//
\glb we \Det.\Pl{} children play.1\Pl{} \Det.\Pl{} cards//
\glft `We children play cards.'\\\citep[276, (23c)]{hoehnetalICGLcalabria}//
\endgl
\xe

See similarly \citet[142]{hoehnetal2016CalabrUnagr}

\subsubsection{Russian (russ1263), Slavic}

\ex \begingl
\gla \textbf{my} leningradcy//
\glb we Leningrader.\Pl//
\glft `we Leningraders'\\\citep[352, (4c)]{pesetsky1978}//
\endgl
\xe

\newpage

\subsubsection{Bulgarian (bulg1262), Slavic}

\ex \begingl
\gla \textbf{nie} student-i-te//
\glb we students-\Pl-\Def//
\glft `we students'\\\citep[560]{hoehn2015unagr}//
\endgl
\xe

See also \citet{osenova2003} and \citet{norman2001} for discussion on unagreement phenomenon in Bulgarian.


\subsubsection{Pomak (poma1238), Slavic}

\ex
\begingl
\gla \textbf{'nami} Po'matsem-se no na p\textturnv{}'maga 'nikutri//
\glb we.\Dat{} Pomak.\Dat.\Pl{}-\Det.\Prox{} 1\Pl.\Acc{} \Neg{} help.3\Sg{} nobody//
\glft `Nobody helps us Pomaks.'\\ \citep[582]{papadimitriou2008}//
\endgl
\xe 

See also \citet[269--272]{hoehnPhD} for some related discussion.

\subsubsection{Polish (poli1260), Slavic}

\ex \begingl
\gla \textbf{My} lingwiści lubimy formalizację.//
\glb we linguists like.1\Pl{} formalisation//
\glft `We linguists like formalisation.'\\\citep[161f., (3b)]{rutkowski2002}//
\endgl
\xe


%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{Khoe-Kwadi languages}

\subsubsection{Khoekhoe/Nama (nama1264), Khoe-Kwadi}


\ex \begingl
\gla \textbf{sa} k\textbottomtiebar{h}oe-\textbf{ta} ké n\~{i} ra \textdoublepipe{}'o.//
\glb \Art{}.\Addr{} person-1\Pl{}.\Incl{}.\Common{} \Top{}? \Compel{} \Prog{} die//
\glft `We humans have to die.' \\{\citep[after][133, (27b)]{boehm1985}}//
\endgl
\xe

Thanks to Menán du Plessis for Khoekhoe glossing. Khoekhoe glosses for \gls{persn}-expressions are my interpretation of \citet{haacke1977}.
See \citet[133--145]{boehm1985} and \citet[140]{maho1998} and particularly \citet{haacke1976, haacke1977, haacke2013namamorph} for further discussion. 

\subsubsection{Khwe/Kxoe (kxoe1243), Khoe-Kwadi}

\ex
\begingl
\gla {Hè é} \textbf{tó} Khwé-\textbf{tò} dì gó\'{\textepsilon} à tó ò + \textbf{\textdoublepipe{}é} Qúva-\textbf{\textdoublepipe{}è} \textdoublebarpipe{}xà-á-tè à.//
\glb \Dem{} 2\Pl{}.\Common{} Khwe-2\Pl{}.\Common{} \Poss{} cattle \Obj{} 2\Pl{}.\Common{} \Poss{} 1\Pl{}.\M{} White-1\Pl{}.\M{} give-I-\Prs{} \Obj{}//
\glft `Here are yours, the Khwe's cows that we, the Whites, give you.' \\{(\citealp[41, (1)]{kilianhatz2008} quoting \citealp[514f.]{koehler1989})}//
\endgl
\xe

See \citet[ch. 3.1.2]{kilianhatz2008} for discussion and \citet[79]{kilianhatz2008} for a further example.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{Niger-Congo languages}


\subsubsection{Kinyarwanda (kiny1244), Bantu}


\ex \begingl
\gla N-ubah-a \textbf{wow} mu-kire.//
\glb 1\Sg.\Subj-respect-\Fv{} 2\Sg{} \Ncl{}1-rich.man//
\glft `I respect you rich man.' \\\citep[69, (73a)]{vanderwal2022}//
\endgl
\xe

See \citet[67--69]{vanderwal2022} for discussion.

\subsubsection{Luganda (gand1255), Bantu}

\ex \begingl
\gla \textbf{Ffe} abantu abaavu ffe tubonaabona.//
\glb we people poor 1\Pl{} suffer//
\glft `We poor people suffer.'\\(elicited)//
\endgl
\xe 

See also \citet[102]{ashtonetal1954} for further (unglossed) examples.

\newpage

\subsubsection{Nkore-Kiga (nkor1241), Bantu}

\ex 
\begingl
\gla \textbf{itwe} abanyankore ni-tu-hinga ebinyoobwa//
\glb we Banyankole \Prs.\Ipfv-1\Pl-cultivate groundnuts//
\glft `We Banyankole grow groundnuts.'\\\citep[after][131, (368)]{taylor1985}//
\endgl
\xe 


\subsubsection{Nzadi (nzad1234), Bantu}


\ex \begingl
\gla nt\^{\textschwa}m. y\textepsilon{} kó luzí\ng{} \textdownstep{}é \textbf{b\"{i}} andzéé b\v{i} k yéè ntswé ninyá b\textopeno{}. atá{[\ldots]}//
\glb taste and \Loc{} life of us Nzadi.people we \Neg.\Prs{} sell fish that \Neg{} even//
\glft `\ldots tasty. In the tradition of us Nzadi, we don't sell that fish. Even\ldots' \\\citep[279, (10)]{craneetal2011}//
\endgl
\xe 

See \citet[210, 279]{craneetal2011}.

\subsubsection{Swahili (swah1253), Bantu}


\ex 
\begingl
\gla \textbf{Nyinyi} wa-nafunzi m-me-cheka.//
\glb you.\Pl{} \Ncl{}2-student 2\Pl{}-\Pst-laugh//
\glft `You students laughed.'\\(elicited)//
\endgl
\xe

Concerning unagreement \citep{hurtado1985, ackemaneeleman2012unagr} in Swahili see \citet[546]{hoehn2015unagr} for a short note.

\subsubsection{Babungo (veng1238), Wide Grassfields}

\ex
\begingl
\gla  \textbf{yìa} v\'{\textbari}\textbari{} ndâa g\'{\textschwa} ntó'//
\glb we.\Excl{} people smithy go-\Prs{} palace//
\glft `We, the blacksmiths, go to the palace.'\\\citep[197, (134a)]{schaub1985}//
\endgl
\xe

See \citet[197f.]{schaub1985} for discussion.

\subsubsection{Koromfe (koro1298), Koromfe}

\ex 
\begingl
\gla \textbf{\textupsilon{}k\textopeno{}} (a) koromb\textturnv{}//
\glb we \Art{} Koromba//
\glft `we Koromba'\\\citep[251, (585)]{rennison1997})//
\endgl
\xe 

The article may be dropped in fast speech, for more discussion see  \citet[242, 250f.]{rennison1997}.

\subsubsection{Supyire (supy1237), Senufo}

\ex \begingl
\gla \textbf{wùu} shìin taanré//
\glb we person.\Ncl{}1.\Pl{} three//
\glft `we three'\\\citep[after][208, (45a)]{carlson1994}//
\endgl
\xe

See \citealp[207f.]{carlson1994}.


%%%%%%%%%%%%%%%%%%%%%%
\subsection{``Papuan'' languages (non-Austronesian languages of Oceania)}


\subsubsection{Menya (meny1245), Nuclear Angan}

\pex \label{ex:menyabpc}
\a \begingl
\gla Nyi tä=\ng{}ga=\ng{}i Matiu i=qu=k=\textbf{i} kuk\ng{}uä + hn=i yat\ng{}qä k-i-m=\ng{}qä=i.//
\glb 1\Sg{} this=\textsc{time}=\Gvn{} Matthew \Dem=\M={2\Sg}=\Obj{} talk \Indf=\F{} ask 2\Sg-do-1\Sg/\Irr=\textsc{goal}=\Ind{}//
\glft `I'm now going to ask you Matthew something.'//
\endgl
\a\begingl
\gla \textbf{Ne} ämaqä qokä i=qu=\textbf{ne} yiämisa\ng{}ä huiyi=nä qw ä-n-k-qäqu=i.//
\glb 1\Pl{} person man \Dem=\M={1\Pl} food other=\Foc{} \Cert{} \Ass-eat-\Pst/\Pfv{}-1\Pl/\Dso=\Ind{}//
\glft `Then we men ate some other food.'\\
\citep[30, (58/59)]{whitehead2006}//
\endgl
\xe

Possible person marking on an indefinite expression:

\ex \begingl
\gla Hn=qu=\textbf{ki} quwä ä-ma-t-qä=i//
\glb \Indf{}=\M=\textbf{2\Sg} steal \Ass-get-2\Sg/\Irr-\Generic=\Def{}//
\glft `Should one of you steal something\ldots'\\\citep[9, (18)]{whitehead2013}//
\endgl
\xe

See \citet[40, 46, 56f.]{whitehead2006} and \citet[9f., 18f.]{whitehead2013} for further examples and discussion.

\subsubsection{Momu (fass1245), Baibai-Fas}

\ex 
\begingl
\gla \textbf{Yery} mu, baso nenwu wu-ta-r-u.//
\glb 1\Pl{} women child belly \Inan:be.at-\Stvzr-1\Pl\Sg{}-\Nmlz//
\glft `For we women, we get pregnant (lit. children are in our bellies).'\\\citep[568, (4)]{honeyman2016}//
\endgl
\xe

See \citet[169f., 242, 568]{honeyman2016}.

\subsubsection{Imonda (imon1245), Border}


\ex
\begingl
\gla \textbf{ka} sebuhe t\~{o}g\~{o} fi-li-t//
\glb 1 devil thus do-\Emph-\Cf//
\glft `We devils should have done it like that.'\\\citep[61, (8)]{seiler1985}//
\endgl
\xe 

See \citet[44, 61f.]{seiler1985}. The possibility for postnominal \glspl{apc} is mentioned, but unfortunately no example is provided.

\subsubsection{Bilua (bilu1245), Bilua}

\pex
\a
\begingl 
\gla \textbf{enge}=a Solomoni=a=ma maba poso=\textbf{ngela}//
\glb 1\Pl{}.\Excl{}=\Lig{} Solomon=\Lig=3\Sg{}.\F{} person \Pl{}.\M{}=1\Pl{}.\Excl{}//
\glft `we, Solomon people' \\{\citep[85, (7.35)]{obata2003}}//
\endgl
\a 
\begingl
\gla \ldots{} lai=za=\textbf{mu}=\textbf{mela} inio me.//
\glb {} where=\Lig=3\Pl{}=2\Pl{}{} \Foc{}.\Nf{} 2\Pl{}//
\glft `\ldots you are people from where?' \\\citep[88, (7.49)]{obata2003}//
\endgl
\xe

See \citet[47--49, 76, 79, 84f., 87--89, 92f.]{obata2003} for a range of further examples.

\subsubsection{Lavukaleve (lavu1241), Lavukaleve}

\ex 
\begingl
\gla aka {malav} \textbf{e} roa-ru kiu-la-m.//
\glb then people 1\Pl{}.\Excl{} one.\Sg{}\M{}-none die-\Neg{}-\Sg{}\M{}//
\glft `And we, the people [lit: the people we] didn't die. [i.e. None of us people died.]'\\{\citep[after][171, (197)]{terrill2003}}//
\endgl
\xe

See \citet[171--173]{terrill2003} and \citet[435, (27)]{terrill2004}.

\subsubsection{Moskona (mosk1236), East Bird's Head}

\pex
\a
\begingl
\gla \textbf{mi}-osnok mi-en-ah-miy,  mi-en-ot jig miyes + mi-er tofi.//
\glb 1\Pl{}-person 1\Pl-\Dur{}-bathe 1\Pl-\Dur{}-stand \Loc{} clothes 1\Pl{}-wear hat//
\glft `we people bathe, wear clothes, wear hats\ldots' (stand in clothes = wear clothes)\\
\citep[344, (2)]{gravelle2010}//
\endgl
\a
\begingl
\gla \textbf{Eri} \textbf{i}-ejen(a) i-odog jig\ldots//
\glb they.\Pl{} 3\Pl-woman 3\Pl-pregnant \Loc{}//
\glft `(If/when) they the women are pregnant\ldots' \\
\citep[91, (38)]{gravelle2010}//
\endgl
\a 
\begingl
\gla \textbf{eri} Mosmir//
\glb they Maybrat//
\glft `Maybrat people/tribe'\\
\citep[224, (43b)]{gravelle2010}//
\endgl
\xe

See \citet[91, 222--224, 344]{gravelle2010}.
 
\subsubsection{Sougb (mani1235), East Bird's Head}

\ex 
\begingl
\gla \textbf{Emen} Sougb/ emen meijouhw dangga.//
\glb we.\Excl{} Sougb we.\Excl{} custom like.that//
\glft `We Sougb, that's our custom.' \\\citep[274, (40)]{reesink2002sougb}//
\endgl
\xe 

See \citet[200, 269f., 274]{reesink2002sougb}.

\subsubsection{Hatam (hata1243), Hatim-Mansim}

\ex \begingl
\gla \textbf{Nye}-ni sop Tinam hwop yok nya ci gi-mbres//
\glb we-this woman Tinam girl put \Pl{} chase-away \Nmlz{}-wide//
\glft `We Tinam women are not easy to take.'\\\citep[195, (72)]{reesink1999}//
\endgl
\xe 


\subsubsection{Maybrat (maib1239), Maybrat}

\ex
\begingl
\gla fnia \textbf{anu} p-no po re-t-o fawen fe//
\glb woman 2\Pl/1\Pl.\Incl{} 1\Pl-do thing location.\Spec-\Prox-\Unmark{} long.time \Neg{}//
\glft `We women (inc), we haven't done this thing for a long time.'\\\citep[158, (99)]{dol2007}//
\endgl
\xe

See \citet[64, fn. 5]{dol2007} concerning the 1\Pl.\Incl{} use of the pronoun \emph{anu}.
See \citet[141, 158, 172, (281)]{dol2007} for examples.

\subsubsection{Urim (urim1252), Urim}

\ex \begingl
\gla \textbf{men} melnum watipmen//
\glb 1\Pl.\Excl{} person many//
\glft `we(excl.) many people'\\
\citep[123]{hemmilaeluoma1987}//
\endgl
\xe 

\pex
\a
\begingl
\gla \textbf{tu} melnum//
\glb 3\Pl{} person//
\glft `the people'//
\endgl
\a \begingl
\gla \textbf{kupm} Melming la nak-etn por ti//
\glb 1\Sg{} Melming say tell.\Real{}-2\Sg.\Obj{} story this//
\glft `I Melming told you this story.'\\
\citep[125]{hemmilaeluoma1987}//
\endgl
\xe

See \citet[123, 125]{hemmilaeluoma1987}.

\subsubsection{Savosavo (savo1255), Savosavo}

\ex \begingl
\gla {}[[No \textbf{mapa=gha}]\unt{NP} [\textbf{ave}]\unt{NP}]\unt{NP}=na kula ata no-va nito=la.//
\glb 2\Sg.[\Gen] person=\Pl{} 1\Pl.\Excl=\Nom{} seawards here 2\Sg-\Gen.\M{} eye=\Loc.\M{}//
\glft `[Addressing the volcano:] We, your people, (are) here seawards at your eye.'\\\citep[155, (285)]{wegener2012}//
\endgl
\xe

See \citet[147,155--159]{wegener2012}.

\subsubsection{Manambu (mana1298), Sepik, Ndu}

\ex \begingl
\gla \textbf{wun} ñ\textschwa{}n-a ñam\textschwa{}s//
\glb I you.\F-\Lnk+\F.\Sg{} younger.sibling//
\glft `me, your younger sister'\\\citep[197]{aikhenvald2008}//
\endgl
\xe

See \citet[197f., 508--513]{aikhenvald2008} for further examples and discussion.

\subsubsection{Awtuw (awtu1239), Sepik, Ram}


\pex \a
\begingl
\gla \textbf{rom} Meley-y\ae{}nim//
\glb 3\Pl{} Meley-\Generic{}//
\glft `the people from Meley' \\\citep[122, (15a)]{feldman1986}//
\endgl
\a 
\begingl
\gla \textbf{rey}/ \textbf{tey} tale//
\glb 3\M.\Sg{}/ 3\F.\Sg{} woman//
\glft `the woman'\\\citep[123, (21a)]{feldman1986}//
\endgl
\xe

See \citet[120--124]{feldman1986} for discussion.

\newpage

\subsubsection{Alamblak (alam1246), Sepik, Sepik Hill}

\pex %\citet[96]{bruce1984} \trailingcitation{[Alamblak]}
\vspace{-1.832\baselineskip}
\begin{multicols}{3}
\a
\begingl 
\gla yima-\textbf{m}//
\glb person-3\Pl{}//
\glft `people'\\\citep[96]{bruce1984}//
\endgl
\a
\begingl
\gla yima-\textbf{k\"{e}}//
\glb person-2\Pl{}//
\glft `you people'//
\endgl
\a
\begingl
\gla yima-\textbf{n\"em}//
\glb person-1\Pl{}//
\glft `we people'//
\endgl
\end{multicols}
\xe

See \citet[90--92, 96f.]{bruce1984}.

\subsubsection{Fore (fore1270), \gls{tng}, Fore-Gimi}


\ex
\begingl
%\gla aogi yagaraná: kanauwe.//
\gla aogi yagara:'-\textbf{na:} kana-u-e//
\glb good man-1\Sg{} come-1\Sg-\Ind{}//
\glft `I, the good man, come.'\\{\citep[after][80]{scott1978}}//
\endgl
\xe

See \citet[79f., 100f.]{scott1978}.

\subsubsection{Hua (huaa1250), \gls{tng}, Siane-Yagaria}

\pex 
\a
\emph{Forapi' + da} $\rightarrow$ /\emph{forapi \textbf{da}}/ `I, Forapi'
\a
\emph{Forapi' + Ka} $\rightarrow$ /\emph{forapi\textbf{ga}}/ `You, Forapi'
\a
\emph{nono' + 'Kama' + da} $\rightarrow$ /\emph{nonokama \textbf{da}}/ `I your maternal uncle'\\{\citep[226]{haiman1980}}
\xe

\pex Person marked genitive forms{\citep[after][240]{haiman1980}}
\a vimata \newline`of us men'
\a ademata \newline `of us women'
\a vi'ita \newline `of you men'
\a adita \newline `of you women'
\xe

See \citet[226--232, 239f.]{haiman1980}.

\subsubsection{Yagaria (yaga1260), \gls{tng}, Siane-Yagaria}

\ex 
\begingl
\gla dagaea ve \textbf{agaea} $\emptyset$-begi-d-u-e//
\glb I man he him-hit-\Pst{}-1\Sg{}-\Ind{}//
\glft `I hit the man.' \\{\citep[18f.]{renck1975}}//
\endgl
\xe

\pex\a
\begingl
\gla Ovu-\textbf{da} ma-lo' bei-d-u-e//
\glb Ovi-I this-\Loc{} live-\Pst{}-1.\Sg{}-\Ind{}//
\glft `I, Ovu, am here.'//
\endgl
\a
\begingl
\gla a-\textbf{tata} e-d-a'-e//
\glb woman-they.\Du{}{} come-\Pst{}-3.\Du{}-\Ind{}//
\glft `The two women came.'\\{\citep[19]{renck1975}}//
\endgl
\xe

See \citet[17f., 166, 181]{renck1975}.

\subsubsection{Amele (amel1241), \gls{tng}, Mabuso}

\ex \begingl
\gla {}[{Dana} {ben} {eu} \textbf{age}] ho-ig-a.//
\glb man big that 3\Pl{} come-3\Pl{}-\Todpst{}//
\glft `Those leaders (big men) came.'\\{\citep[after][210, (283)-(284)]{roberts1987amele}}//
\endgl
\xe

See \citet[162, 201, 209f.]{roberts1987amele}.

\subsubsection{Kobon (kobo1249), \gls{tng}, Kalam-Kobon}

\pex
\a
\begingl
\gla Juab Minöp \textbf{kal\textbari{}p} ñ-öb.//
\glb Juab Minöp \Obj.3\Du{}{} give-\Prf.3\Sg{}//
\glft `He gave it to Juap and Minöp.'\\{\citep[after][108, (264)]{davies1989}}//
\endgl
\a
\begingl
\gla \textbf{Yad} Kaunsol nibi b\textbari{} abad aij g\textbari{}-m\textbari{}d-pin.//
\glb 1\Sg{} councillor woman man look.after good do-\Habit-\Pfv.1\Sg{}//
\glft `As councillor I look after the people well.' \\\citep[157, (408ad)]{davies1989}//
\endgl
\xe 

See  \citet[107f., 157]{davies1989} for instances of prenominal pronouns as well as postnominal ``pronominal copies''.

\subsubsection{Usan (usan1239), \gls{tng}, North Adelbert} 

\pex
\a \begingl
\gla eâb igim-ine ne tain \textbf{wo} yâ-nâmb wogub\ldots//
\glb cry.\Ss{} be-1\Sg.\Ds{} and father he me-hit.\Ss{} cease.\Ss{}//
\glft `I was crying and my father he hit me and then\ldots' \\
\citep[after][167, (99)]{reesink1987}//
\endgl
\a 
\begingl
\gla \ldots{} in bo \textbf{an} wau moi qomon gâs ende ig-oun//
\glb {} we again you child unmarried custom like thus be-1\Pl.\Prs{}//
\glft `\ldots we in turn live like the customs of you young men.'\\
\citep[part of][190f., (41)]{reesink1987}//
\endgl
\xe 

See \citet[53f., 167, 190f., 353]{reesink1987}.


\subsubsection{Adang (adan1251), \gls{tng}, Alor-Pantar}

\ex 
\begingl
\gla Sa [Bain \textbf{\textglotstop{}ari}]\unt{NP} b$\epsilon$h.//
\glb 3\Sg.\Sbj{} Bain 3.\Obj{} hit//
\glft `S/he hit Bain.'\\\citep[261, (165)]{robinsonhaan2014}//
\endgl
\xe

\subsubsection{Kaera (kaer1234), \gls{tng}, Alor-Pantar}

\ex 
\begingl
\gla Ilwang \textbf{gang} user\texttildelow{}user bir bleling g-om mi eser-o.//
\glb Ilwang 3\Sg{} \Redup\texttildelow{}quickly run open 3\Sg{}-inside \Loc{} exit-\Fin{}//
\glft `Ilwang quickly ran outside.' (lit. `\ldots ran out to (the) open's inside')\\\citep[129, (98)]{klamer2014}//
\endgl
\xe

\newpage

\subsubsection{Kamang (kama1365), \gls{tng}, Alor-Pantar}

\ex
\begingl 
\gla almakang=ak \textbf{gera}//
\glb people=\Def{} 3.\Contr{}//
\glft `the \{specific group of\} people'\\{\citep[313f., (58a)]{schapper2014}}//
\endgl
\xe

See \citet[313f.]{schapper2014}.

\subsubsection{Sawila (sawi1256), \gls{tng}, Alor-Pantar}

\ex \begingl
\gla {}[aning du \textbf{girra}]\unt{A} [parra]\unt{P} laata//
\glb \Nfin.person \Pl{} 3.\Aarg{} field burn//
\glft `People are burning fields.'\\ \citep[392, (102f)]{kratochvil2014}//
\endgl
\xe

See \citealp[391f.]{kratochvil2014}.
 
\subsubsection{Wersing (wers1238), \gls{tng}, Alor-Pantar}

\ex \begingl
\gla aning \textbf{gnuk} unan le-wena//
\glb person 3.\Du{} louse \Appl-search//
\glft `Those two people searched for lice.'\\\citep[472, (75)]{schapperhendery2014}//
\endgl
\xe

See \citet[472]{schapperhendery2014}.


\subsubsection{Western Pantar (lamm1241), \gls{tng}, Alor-Pantar}

\ex 
\begingl
\gla {}[[Tabang alaku Duinni Maggangkala]\unt{NP} [\textbf{ging}]\unt{NP}]\unt{NP} a-raung yattu ga-ung misingup.//
\glb slave two Duinni Maggangkala 3\Pl.\Act{} INCP-climb tree 3\Sg-head sit//
\glft `The two slaves Duinni Maggangkala climbed up and sat in the tree.'\\\citep[53, (105)]{holton2014}//
\endgl
\xe 

\citet[53, fn. 3]{holton2014} notes that ``Duinni Maggagkala is a single (binomial) name give [sic] to the pair of slaves together.''
Further see \citet[53f.]{holton2014}.


%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{Sino-Tibetan languages}


\subsubsection{East Geshiza (gesh1238), Burmo-Qiangic} 

\ex 
\begingl
\gla rd\textctz{}\ae{} \textbf{lm\ae{}=\textltailn{}\textschwa}=t\super{h}\textschwa~t\super{h}\textschwa{} mp\super{h}ri v-s\super{h}\ae{}=b\textopeno{}, rd\textctz{}\ae{}. b\ae{} \ng\ae{}=\textltailn{}\textschwa{}=t\super{h}\textschwa{} mp\super{h}ri mi-s\super{h}o\ng{}.//
\glb Chinese 3=\Pl.\Abs{}=\Top~\Redup{} snake.\Abs{} \Inv-kill.\Npst.3=\Mod{} Chinese.\Abs{} Tibetan 1=\Pl.\Abs{}=\Top{} snake.\Abs{} \Neg{}-kill.\Npst.1\Pl{}//
\glft `The (Han) Chinese, they kill snakes. [\ldots] We Tibetans do not kill snakes.'\\\citep[480, (7.48)]{honkasalo2019}//
\endgl
\xe

See \citet[388, 400, 438, 480, 507, 646]{honkasalo2019} for further examples.

\subsubsection{Mandarin (mand1415), Chinese}


\ex \begingl
\gla congming de \textbf{ni-men} ziji xiang banfa jiejue ba!//
\glb smart \Lnk{} you-\Pl{} yourselves think ways solve \Exclam{}//
\glft `you smart people think of a way to solve it yourselves.'\\\citep[after][(53a)]{boskovichsieh2013}//
\endgl
\xe 

See \citet[sec. 7.3]{boskovichsieh2013} for further examples.


%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{Uralic languages}


\subsubsection{Hungarian (hung1274), Ugric}

\ex \begingl
\gla \textbf{Ti} orvos-ok sok-at dolgoz-tok.//
\glb you.\Pl{} doctor-\Pl{} much-\Acc{} work-\Indf.2\Pl//
\glft `You doctors work a lot.' \\\citep[269, (492)]{keneseietal1998}//
\endgl
\xe

See \citet[269]{keneseietal1998} and \citealp[559]{hoehn2015unagr}.

\newpage

\subsubsection{Finnish (finn1318), Finnic}

\ex \begingl
\gla \textbf{Me\textbf{}} naiset menemme nyt saunaan.//
\glb we woman.\Pl{} go.1\Pl{} now sauna.\Ill//
\glft `We women will go to the sauna now.'\\\citep[277, (1335)]{sulkalakarjalainen1992}//
\endgl
\xe

See also \citet[24f.]{hoehn2020ThirdGap}.


%%%%%%%%%%%%%%%%%%%%%%%
\subsection{Creoles}


\subsubsection{Ndyuka (ndyu1242), Creoles}

\ex \begingl
\gla \textbf{u} gaanman fu den liba//
\glb 1/2\Pl{} chief for the.\Pl{} river//
\glft `you chiefs of the rivers'\\\citep[467, (2075)]{huttarhuttar1994}//
\endgl
\xe

See \citet[224, 460, 466f.]{huttarhuttar1994}.

\subsubsection{Nigerian Pidgin (nige1257), Creoles}

\ex \begingl
\gla \textbf{Unà} onyìbo pipul no dè chu kola àt\b{ô}l.//
\glb you.\Pl{} white people \Neg{} \Ipfv{} chew kola \Neg.\Emph{}//
\glft `You white people don't chew kola nut at all.'\\\citep[after][181, (802)]{faraclas1996}//
\endgl
\xe

See \citet[178, 181]{faraclas1996}.

\subsubsection{Kristang (mala1533), Creoles}

\ex
\begingl
\gla  kora jenti muré, tudu \textbf{nus} kristáng bai//
\glb when person die all 1\Pl{} {} go//
\glft `When people die, all we Kristangs go (to the wake).' \\\citep[86, (11)]{baxter1988}//
\endgl
\xe

See \citet[86]{baxter1988}.


%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{Various}


\subsubsection{Mi'kmaq (mikm1235), Algonquian}

\ex
\begingl
\gla \textbf{ninen} elnui-yek//
\glb we.\Excl{} people-1\Pl.\Excl//
\glft `we First Nation people'\\
\citep[188]{pacifiqueetal1990}//
\endgl
\xe

Thanks to Watson Williams for help with the glossing.
See \citet[188]{pacifiqueetal1990} for some further (unglossed) examples.

\subsubsection{Katu (nucl1297), Katuic}

\pex
\a \begingl
\gla manuih \textbf{yi}//
\glb people we//
\glft `we people'//
\endgl
\a 
\begingl
\gla \textbf{yi} manuih//
\glb we people//
\glft `we people'//
\endgl
\a
\begingl
\gla \textbf{yi} adi anó \textbf{yi}//
\glb we older.brother younger.brother we//
\glft `we older and younger brothers'\\
\citep[28, (35--37)]{costello1969}//
\endgl
\xe

See \citet[28]{costello1969}.

\subsubsection{Wari' (wari1268), Chapacura-Wanham}

\ex \begingl
\gla \textbf{Wirico} Mon' co pa' na mijac//
\glb \Emph:3\Sg.\M{} \M:name \Infl:\M/\F.\Real.\Pst/\Prs{} kill 3\Sg:\Real.\Pst/\Prs{} pig//
\glft `It was Mon' who killed a pig.'\\\citep[after][303, (570a)]{everettkern1997}//
\endgl
\xe

See \citet[303, 310]{everettkern1997}.

\newpage

\subsubsection{Chitimacha (chit1248), Chitimacha}

\ex \textbf{\textglotstop{}u\v{s}} pan\v{s}' ha hananki' namkinada'\\
`We people who live in this house.'\\
\citep[333]{swadesh1967}\xe

See \citet[333]{swadesh1967}.

\subsubsection{Lezgian (lezg1247), Lezgic}

\ex 
\begingl
\gla \textbf{\v{C}a-z} \v{c}uban-r.i-z, wun har näni-q$^h$ k'wal.i-z q$^h$ifi-zwazj-di q$^h$sam \v{c}i-zwa//
\glb we-\Dat{} shepherd-\Dat{} you:\Abs{} every night-\Postess{} house-\Dat{} return-\Ipfv-\Ptcp-\Nmlz good know-\Ipfv//
\glft `We shepherds know well that you go home every night.' \\\citep[after][259, (682a)]{haspelmath1993}//
\endgl
\xe

See \citet[259]{haspelmath1993}.

\subsubsection{Abkhaz (abkh1244), Northwest Caucasian}

\ex 
\begingl
\gla \textbf{\textcrh{}a(rà)} (\textbf{\v{s}$^o$a(rà)}, \textbf{darà}) a-bà\textcrh\v{c}-aa-ja-y$^o$-c$^o$a//
\glb we you they \Art-garden-\Prev-tend-\Aarg{}-\Pl//
\glft `we (you, they) gardeners' \\\citep[159]{hewitt1989}//
\endgl
\xe

See \citet[157, 159]{hewitt1989}.

\subsubsection{Basque (basq1248), Basque}

\pex
\a
\begingl
\gla Galdu didazue aita-seme-\textbf{ok} afari-ta-ko gogo guzti-a.//
\glb spoil 3\Sg{}.\Abs{}.\Aux{}.1\Sg{}.\Dat{}.\textbf{2\Pl{}.\Erg} father-son-\Proxart.\Pl{}.\Erg{} dinner-\Loc{}-\Lnk{} appetite all-\Det{}.\Abs{}//
\glft `{You, father and son}, have spoiled my whole appetite for dinner.' \\{\citep[502, (90a)]{derijk2008}}//
\endgl 
\a
\begingl
\gla Zor berri-a dugu euskaldun-\textbf{ok} Orixe-rekin.//
\glb debt new-\Det{}.\Abs{}{} 3\Sg{}.\Abs{}.\Aux{}.\textbf{1\Pl{}.\Erg} Basque-\Proxart.\Pl{}{} Orixe-\Com{}//
\glft `{We Basques} have a new debt to Orixe.' \\{\citep[502, (91a)]{derijk2008}}//
\endgl
\xe

See \citet[210]{saltarelli1988}, \citet[122]{trask2003}, \citet[482, 501f.]{derijk2008}, \citet[67]{areta2009} and especially \citet{artiagoitia2012DP}.

\subsubsection{Classical Nahuatl (clas1250), Aztecan}

\pex
\a
\begingl
\gla \textbf{Ni}cu\={\i}ca \textbf{ni}Petoloh.//
\glb I-sing I-am-Peter//
\glft `I, Peter, sing.'//
\endgl
\a
\begingl
\gla \textbf{N\={e}ch}itta \textbf{ni}Petoloh.//
\glb he-sees-me I-am-Peter//
\glft `He sees me, Peter.'\\{\citep[193]{andrews1975}}//
\endgl
\a
\begingl
\gla \textbf{No}cal \textbf{ni}Petoloh.//
\glb It-is-my-house/they-are-my-houses I-am-Peter//
\glft `It is my house (and I am Peter)./They are my houses (and I am Peter).'\\{\citep[194]{andrews1975}}//
\endgl
\xe

See \citet[192--194]{andrews1975} and \citealp[ch. 17.3]{andrews2003}.

\subsubsection{Evenki (even1259), Tungusic}

\ex \begingl
\gla  \textbf{Bu} bejumimni-l eme-re-$\emptyset$.//
\glb we hunter-\Pl{} come-\Nfut-3\Pl//
\glft `We, hunters, came.' \\\citep[199, (794)]{nedjalkov1997}//
\endgl
\xe

See \citet[197, 199]{nedjalkov1997}.



\subsubsection{Hixkaryana (hixk1239), Cariban}

\citet[131]{derbyshire1979} suggests that Hixkaryana does not have integrated adnominal person marking (and no adnominal demonstratives), but instead paratactic constructions like (\nextx).

\pex
\a
\begingl
\gla m\textbari{}nayar\textbari{} hor\textbari{} \textbf{amna} ntono. n\textbari{}mno + hokono rma \textbf{amna}//
\glb species.of.leaf seeking we.\Excl{} went house {one.occupied.with} same-ref we.\Excl{}//
\glft `We housebuilders went looking for leaves'\\{\citep[131, (290)]{derbyshire1979}}//
\endgl
\a
\begingl
\gla nux mokro raheno//
\glb my.younger.brother that.one he.seduced.me//
\glft `That younger brother of mine seduced me'\\{\citep[132, (293a)]{derbyshire1979}}//
\endgl
\xe



\subsubsection{Japanese (nucl1643), Japanese} 

\ex \begingl
\gla \textbf{wareware} nihonjin//
\glb we Japanese//
\glft `we Japanese'\\\citep[254]{hinds1988}//
\endgl
\xe

See \citet[254, 261]{hinds1988} and for more detailed discussion with different theoretical proposals: \citet[780]{noguchi1997}, \citet[sec. 3.2]{furuya2008} and \citet{inokuma2009}.

\subsubsection{Korean (kore1280), Korean}


\ex \textbf{wuli-(tul)} hankwuk salam\\
 `we Koreans'\\
 \citep[292]{sohn1994}
 \xe
 
\pex
\a \begingl
\gla \textbf{wuli} ttokttokhan enehakcatul//
\glb we smart linguists//
\endgl
\a \begingl
\gla ttokttokhan \textbf{wuli} enehakcatul//
\glb smart we linguists//
\glft `we smart linguists'\\\citep[151, (15)]{choi2014phd}//
\endgl
\xe

See \citet[284; 292]{sohn1994} and \citet[151--154]{choi2014phd}.


\subsubsection{Kalaallisut/West Greenlandic (kala1399), Eskimo}


\pex
\a
\begingl
\gla kalaalli-t \textbf{uagut}//
\glb Greenlander-\Abs{}.\Pl{} we//
\glft `we Greenlanders'\\{\citep[after][110; gloss extrapolated]{fortescue1984}}//
\endgl
\a
\begingl
\gla \textbf{uagut} kalaali-u-sugut//
\glb we Greenlander-be-1\Pl{}.\Ptcp{}//
\glft `we Greenlanders' \\{\citep[after][257]{fortescue1984}}//
\endgl
\xe

See \citet[110, 253, 256f.]{fortescue1984}.

\subsubsection{Lakkia (lakk1238), Kadai}

\ex 
\begingl
\gla \textbf{tau$^{51}$} hou$^{24}$ {\textglotstop{}at$^{55}$ jen$^{11}$ kjã:u$^{24}$}//
\glb 1\Pl{} two sister//
\glft `we, the two sisters'\\\citep[137, (40)]{fan2019}//
\endgl
\xe

\subsubsection{Turkish (nucl1301), Turkic}

\ex 
\begingl
\gla \textbf{biz} Türk-ler vatan-{\i}m{\i}z-{\i} sev-er-iz//
\glb we Turk-\Pl{} mother.country-1\Pl-\Acc{} love-\Aor-1\Pl{}//
\glft `We Turks love our country'\\\citep[297, (1075)]{kornfilt1997}//
\endgl
\xe

See \citet[288, 297f.]{kornfilt1997}.



